---
title: "igraph in R"
author: "Mouwei Lin (lm3756) and Linhao Yu (ly2590)"
date: "`r Sys.Date()`"
output: 
  html_document:
    fig_width: 10
    fig_height: 7.5
---

```{r setup, include=FALSE}
# this prevents package loading message from appearing in the rendered version of your problem set
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


## 1. igraph Basics

First we download and install the package.

```{r}
#install.packages('igraph')  #remove '#' when you first download it.
library(igraph)
```


We will use an open source data set named 'phone.call' from 'navdata'.

```{r}
library(devtools)
#install_github('kassambara/navdata') #remove '#' when you first download it.
library(navdata)
data("phone.call")
head(phone.call)
```
### 1.1 Create igraph Network Object

igraph has its specific network (graph) object called 'igraph'. The simplest way to create a 'igraph' is using graph.formula() to type every node and edge into it.
```{r}
graph.formula(A-B-C-D,A-E-F,Z-D-X)
```

But usually in real world, we have a lot of data so definitely it not a good idea to add each edge manually. More generally, we will use those two methods:

```{r}
library(tidyverse)
# prepare the data
name<-data.frame(c(phone.call$source,phone.call$destination))
nodes<-name%>%
    distinct()%>%
mutate(location=c("western","western","central","nordic","southeastern",
     "southeastern","southeastern","southern","sourthern",
     "western","western","central","central","central","central","central"))
colnames(nodes)<-c("label","location")

edges<-phone.call%>%
    rename(from=source,to=destination,weight=n.call)
```


#### (1) build igraph object from dataframe:
graph_from_data_frame()
To use this method, we need two dataframes, one is edge frame, the other is vertexes frame.

```{r}
net_pc<-graph_from_data_frame(
   d=edges,vertices=nodes,
   directed=TRUE)
```

we can see that the graph is created. We can use V() or E() to visit vertexes or edges, and as_edgelist(net, names=T), as_adjacency_matrix(net, attr="weight") to catch edges list and adjacent matrix:
```{r}
V(net_pc)
V(net_pc)$location
E(net_pc)
as_edgelist(net_pc, names=T) 
as_adjacency_matrix(net_pc, attr="weight")
```


#### (2) build igraph object from adjacent matrix:
graph_from_adjacency_matrix()
If the graph is using 

```{r}
adjacent_matrix<-as_adjacency_matrix(net_pc, attr="weight")
net_am<-graph_from_adjacency_matrix(adjacent_matrix)
```

now we simply plot it to take a look
```{r}
plot(net_pc)
plot(net_am)
```

### 1.2 Basic igraph Visualization Instructions

The plot function in igraph is very strong, it has lots of parameters to make the network graph more beautiful and clearer. In this section we will only give a general introductions to important visualization methods, we will have a more detailed introductions in the next section.

A large number of parameters are used to display various properties of nodes, edges and graphs. The parameters related to nodes start with vertex.XXX, and the parameters related to edges start with edge.XXX 

In addition to specifying the parameters of nodes and edges in plot(), you can also use the previously mentioned V() and E() to add the corresponding properties directly in the igraph object. The difference between the two methods is that the parameters specified in plot() do not change the properties of the plot. For example, we first specify the color of the node according to the position, and the width of the edge according to the weight (these two attributes will be saved in the net_pc object), and then specify the size of the node in the parameters of plot() (proportional to the degree of the node, The node degree is the number of edges connected to this node), the size and position of the node marker, the color of the edge, the size of the arrow, and the degree of curvature of the edge.
```{r}
# Calculate node's degree
deg<-degree(net_pc,mode="all")
# Set up the color
vcolor<-c("orange","red","lightblue","tomato","yellow")
# Set specific node's Color
V(net_pc)$color<-vcolor[factor(V(net_pc)$location)]
# Set specific edge's weight
E(net_pc)$width<-E(net_pc)$weight/2

# Set up vertex.size, vertex.label.cex & dist, edge color & arrow size & curve in graph
plot(net_pc,vertex.size=3*deg,
     vertex.label.cex=.7,vertex.label.dist=1,
     edge.color="gray50",edge.arrow.size=.4, edge.curved=.1)
# Add legend
legend(x=-1.5,y=1.5,levels(factor(V(net_pc)$location)),pch=21,col="#777777",pt.bg=vcolor)
```


### 1.3 Network Layout



